\chapter{Summary of Assertions}
\label{assertions}

\begin{description}
\assert{controller} Controller methods are JAX-RS resource methods annotated with \code{@Controller}.
\assert{all-controllers} All resource methods in a class annotated with \code{@Controller} must
be controllers.
\assert{void-controllers} Controller methods that return void must be annotated with \code{@View}.
\assert{cdi-beans} MVC beans are managed by CDI.
\assert{per-request} Default scope for MVC beans is request scope.
\assert{validation-result} If validation fails, controller methods must still be called if a 
\code{ValidationResult} field or property is defined. 
\assert{event-firing} All events in \code{javax.mvc.event} must be fired. See Javadoc for more
information on each event in that package.
\assert{builtin-engines} Implementations must provide support for JSPs and Facelets.
\assert{extension-engines} CDI beans that implement \code{javax.mvc.engine.ViewEngine} provide
an extension mechanism for view engines.
\assert{selection-algorithm} Implementations must use algorithm in Section \ref{selection_algorithm} 
to select view engines.
\assert{request-forward} Forward request for which no view engine is found.
\assert{exception-wrap} Exceptions thrown during view processing must be wrapped.
\assert{view-resolution} Relative paths to views must be resolved as explained in 
Section \ref{selection_algorithm}.
\assert{null-controllers} The \code{@View} annotation is treated as a default value for any controller 
method that returns a \code{null} value.
\assert{redirect} Support HTTP redirects using the {\tt redirect:} prefix and a controller
return type of {\tt String}.
\assert{annotation-inheritance} Annotation inheritance is derived from \jaxrs\  and extended to 
\mvc\ annotations. 
\assert{csrf-options} CSRF support for configuration options defined by {\tt Csrf.CsrfOptions}.
\assert{csrf-support} CSRF validation required only for controllers annotated by {\tt @POST} 
and consuming the media type {\tt x-www-form-urlencoded}.
\assert{mvc-context} Application-scoped {\tt MvcContext} available for injection and as {\tt mvc} in EL.
\end{description}
